#Testing for correlation between wage bills and league standings of teams in English Premier League


```{r}
install.packages("foreign", repos="http://cran.rstudio.com/")
install.packages("haven", repos="http://cran.rstudio.com/")
install.packages("xlsx", repos="http://cran.rstudio.com/")
install.packages("XML", repos="http://cran.rstudio.com/")

library(foreign)
library(haven)
library(xlsx)



#Testing for correlation between wage bills and league standings of teams in English Premier League
getwd()
library(XML)
junk <- readHTMLTable("http://www.totalsportek.com/money/english-premier-league-wage-bills-club-by-club/", stringsAsFactors=FALSE)
standings2013 <- data.frame(readHTMLTable("http://www.espnfc.us/barclays-premier-league/23/table?season=2013", stringAsFactors= FALSE, header = TRUE))

names(standings2013)[2] <- c("TEAM")
standings2013[2] <- as.vector(standings2013$TEAM)
standings2013 <- na.omit(standings2013)
standings2013 <- standings2013[order(standings2013$TEAM),]
standings2013 <- standings2013[c(1:17,19,20,21),1:24]

EPLwageBills <- data.frame(junk[[1]][-1,-1]) 
names(EPLwageBills) <- c("Club","2013-2014","2014-2015","2015-2016")
EPLwageBills[c(16,17,18,19,20),c(2)] <- c(46,53,54,43,69)
EPLwageBills <- EPLwageBills[order(EPLwageBills$Club),]

test2 <- EPLwageBills[,c(1,2)]
test2[3] <- test2[2]
test2[,1:2] <- standings2013[,2:1]

test2[,c(3)] <- gsub("m","",test2[,c(3)])
test2[,c(3)] <- gsub("£","",test2[,c(3)])
test2[,c(3)] <- as.numeric(test2[,c(3)])

test2014 <- test2

top2014 <- test2014[1:10,c(3)]
bot2014 <- test2014[11:20,c(3)]

var.test(top2014,bot2014)
t.test(top2014,bot2014, alternative = "greater", var.equal = TRUE)
```
#The t-statistic is -2.9608 and the p-value is very high, indicating that the NULL hypothesis is probably True--true difference in means in less than 0. However, the league standards are in a counterintuitive order--finishing with a lower numeral actually means finishing in a higher position. So a negative t-statistic means that the teams with the highest ten wages, earn a positions of lower numerals but of greater success. Therefore, there is strong evidence of a correlation between (suprise) paying your players more, and winning more

#Ran another test for this year's data, just because the league standings seemed remarkably counterintuitive this year. There's still a correlation, but by less.
```{r}
test <- EPLwageBills[,c(1,4)]

test[,c(2)] <- gsub("m","",test[,c(2)])
test[,c(2)] <- gsub("£","",test[,c(2)])
test[,c(2)] <- as.numeric(test[,c(2)])
test$finalPosition <- c(10,5,4,2,8,3,18,11,9,17,7,20,14,6,12,15,1,19,13,16)

top2016 <- test[1:10,c(3)]
bot2016 <- test[11:20,c(3)]

var.test(top2016,bot2016)
t.test(top2016,bot2016, alternative = "greater", var.equal = TRUE)
```


#Let's find out how many goals, on average, were scored off of corner kicks by a team in the 2011-2012 season
```{r}
EPLgoals <- read.csv("https://raw.githubusercontent.com/soccermetrics/project-data/master/english-premier-league-2011-2012/Goals.csv", stringsAsFactors = FALSE)
                 
goalTypes <- unique(EPLgoals$Event)    

##Use tapply to see how many of each type of goal each team had
teamGoals <- tapply(EPLgoals$Event, EPLgoals$Team, table)

##Used sapply to extract the first component (Goals from Corner Kick Crosses) of each list element. Two teams had none so I set their values equal to 0 manually. Then found the mean corner kick goals of the teams
corners <- sapply(teamGoals,function(x) x[1])
corners[c(2,19)] <- 0
mean(corners)
```